CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-09-28 codecheck-paper.tex
go through introduction section

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\end{itemize}

@@

Many areas of scientific research now use computation to either simulate
or analyse their data. These computations are increasingly complex, and
difficult to explain coherently in a paper. (MARWICK). To complement the
traditional route of sharing research by writing papers, there is a
growing trend/demand to share the underlying artifacts, notably code and
datasets, so that others can inspect, reproduce or expand that work
(Figure 1). Some of the earliest proponents were Buckheit and Donoho
(1995) who commented ``advert is scholarship for the paper''. (get
correct quote).

So, assuming that researchers begin to share more artifacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share. The
implementation status of this varies considerably according to the
journal. At one of the spectrum, there are specialist journals that
have been created to accept ``data papers'' (e.g.~\emph{Scientific
  Data}, \textit{Geoscience Data Journal}\ldots{} NAME COUPLE OF
OTHERS). These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request'' whereby authors,
well-intentioned, often cannot provide the data when readers ask for
it.  (ANDREW H: any references to sum up poor/evolving state of data
sharing; Vines et al 2014).

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
their code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in open science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/use
the code, this is worth documenting as early as possible. With this in
mind, we have developed a set of PRINCIPLES and an example workflow that
provides what he hope is a pragmatic way of checking that code works. We
call this system CODECHECK as outlined next.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research.  The left
  half of the diagram shows the diverse range of materials used
  internally within a laboratory.  These materials are often then
  condensed into a paper for sharing with the outside world via the
  research paper, a static PDF document.  Working backwards from the
  PDF to the underlying materials is impossible.  By sharing the
  materials on the left, others outside the lab are able to reproduce
  or build on this work.  (Should we draw a backward arrow showing the
  impossibility of this backward flow?)}
  \label{fig:inverse}
\end{figure}

>>>>

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\end{itemize}

@@

Many areas of scientific research now use computation to either simulate
or analyse their data. These computations are increasingly complex, and
difficult to explain coherently in a paper \citep{marwick_how_2015}. To complement the
traditional route of sharing research by writing papers, there is a
growing trend/demand to share the underlying artifacts, notably code and
datasets, so that others can inspect, reproduce or expand that work
(Figure 1). Some of the earliest proponents were Buckheit and Donoho \cite{buckheit_wavelab_1995} who coined what has been called 
\emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{"An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures."}
%  Donoho (2010) has another paraphrasing of the same quote, which might work better:
% "an article about computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."

So, assuming that researchers begin to share more artifacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share. The
implementation status of this varies considerably according to the
journal. At one of the spectrum, there are specialist journals that
have been created to accept ``data papers'' (e.g.~\href{}{\emph{Scientific
  Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request'' whereby authors,
while possible well-intentioned at the time of writing the article,
often cannot provide the data when readers ask for
it.  (ANDREW H: any references to sum up poor/evolving state of data
sharing; Vines et al 2014).

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
their code cannot be shared, e.g.~``there is no documentation / I do not
want to maintain it / I do not want to give away my code to my
competitors''. Our view, rooted in open science princples, is that
wherever possible sharing code is good for the community, as outlined
ten years ago \cite{Barnes2010-iv}. Having the code freely available, and
suitably archived, provides a valuable resource for others to learn
from, even when it doesn't run or if there is no documentation. However,
with a little effort, we believe that if someone independent can run/execute
the code, this is worth documenting as early as possible. With this in
mind, we have developed a set of PRINCIPLES and an example workflow that
provides what he hope is a pragmatic way of checking that code works. We
call this system CODECHECK as outlined next.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research.  The left
  half of the diagram shows the diverse range of materials used
  internally within a laboratory.  These materials are often then
  condensed into a paper for sharing with the outside world via the
  research paper, a static PDF document.  Working backwards from the
  PDF to the underlying materials is impossible. This probihibits reuse
  and is not only non-transparent for a specific paper, but also 
  ineffective for science as a whole. By sharing the
  materials on the left, others outside the lab are able to reproduce
  or build on this work.}
  \label{fig:inverse}
\end{figure}
